---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE
)
```

# functiondepends

<!-- badges: start -->
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.com/WelcomeToMyVirtualHome/functiondepends.svg?branch=master)](https://travis-ci.com/WelcomeToMyVirtualHome/functiondepends)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/WelcomeToMyVirtualHome/functiondepends?branch=master&svg=true)](https://ci.appveyor.com/project/WelcomeToMyVirtualHome/functiondepends)
[![codecov](https://codecov.io/gh/WelcomeToMyVirtualHome/functiondepends/branch/master/graph/badge.svg)](https://codecov.io/gh/WelcomeToMyVirtualHome/functiondepends)
<!-- badges: end -->

The goal of functiondepends is to allow for tidy exploration of unstructured codebase without evaluation of code. 

## Installation

One can install `functiondepends` from GitHub with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("WelcomeToMyVirtualHome/functiondepends")
```

```{r}
library(functiondepends)
```

## Examples

```{r}
# Create environment for loaded functions 
envir <- new.env()
# Seach recursively current directory
functions <- find_functions(".", envir = envir, recursive = TRUE)
functions
```

Search for dependencies of function `functions_in_path` within parsed functions:

```{r}
dependency <- find_dependencies("find_functions", envir = envir, in_envir = TRUE)
dependency
```

Search for all dependencies of function:

```{r functions_in_path, fig.height=4}
library(ggplot2)
library(dplyr)

dependency <- find_dependencies("find_dependencies", envir = envir, in_envir = FALSE)

dependency %>% 
  ggplot(aes(x = reorder(Source, SourceRep), y = SourceRep, fill = Namespace)) +
  geom_col() +
  coord_flip() +
  labs(title = "functions_in_path", x = "Source")
```

Limit results to functions only parsed from source files (user defined):

```{r target_degree, fig.height=4}
library(ggplot2)
library(dplyr)

dependency <- find_dependencies(unique(functions$Function), envir = envir, in_envir = FALSE)
dependency %>% 
  distinct(Target, TargetInDegree) %>%
  mutate(Target = reorder(Target, TargetInDegree)) %>%
  ggplot(aes(x = Target, y = TargetInDegree)) +
  geom_col() +
  coord_flip() 
```

```{r network, fig.height=4}
library(igraph)
library(dplyr)

edges <- dependency %>% 
  select(Source, Target) %>% 
  filter(!is.na(.))
vertices <- unique(c(edges$Source, edges$Target))

g <- graph_from_data_frame(
  d = edges, 
  vertices = vertices,
  directed = TRUE
)
plot(
  g,
  vertex.color = "grey",
  edge.color = "grey", 
  vertex.size = 20
)
```

